\textbf{SPECFEM3D}\\ \textbf{Cartesian}\\ \textbf{User Manual}

### 2022-12-08 doc/USER_MANUAL/04_creating_databases.tex
user manual updated - minor updates
updates doc

\vspace{1cm}
\noindent {\bf $\bullet$ Sources setup}
\begin{description}
\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a
(tilted) \texttt{FORCESOLUTION} force point source instead of a \texttt{CMTSOLUTION}
moment-tensor source. When the force source does not fall exactly
at a grid point, the solver interpolates the force between grid points
using Lagrange interpolants. This can be useful e.g. for oil industry
foothills simulations in which the source is a vertical force, normal
force, tilted force, or an impact etc. Note that in the \texttt{FORCESOLUTION}
file, you will need to edit the East, North and vertical components
of an arbitrary (not necessarily unitary, the code will normalize it automatically) direction vector of the force vector;
thus refer to Appendix~\ref{cha:Coordinates} for the orientation
of the reference frame. This vector is made unitary internally in
the solver and thus only its direction matters here; its norm is ignored
and the norm of the force used is the factor force source times the
source time function.

When using this option, by default the code can locate the force source
anywhere between mesh points in order to honor its exact location; this
is more precise than using the closest GLL mesh point, but it is also a bit slower.
If needed, you can change that default behavior and force the code to use the closest
GLL mesh point instead by setting flag \texttt{USE\_BEST\_LOCATION} to \texttt{.false.} instead of \texttt{.true.} in file
\texttt{setup/constants.h} and recompiling the code.
\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
a Ricker source time function, i.e., the second derivative of a Gaussian, instead of the source time functions
set by default to represent a (tilted) \texttt{FORCESOLUTION} force
point source or a \texttt{CMTSOLUTION} moment-tensor source.
Note that we use the standard definition of a Ricker, for a dominant frequency $f_0$:
$\mathrm{Ricker}(t) = (1 - 2 a t^2) e^{-a t^2}$, with $a = \pi^2 f_0^2$,
whose Fourier transform is thus:
$\frac{1}{2} \frac{\sqrt{\pi}\omega^2}{a^{3/2}}e^{-\frac{\omega^2}{4 a}}$
This gives the wavelet of Figure~\ref{fig:RickerWavelet}.
%%
\begin{figure}[htbp]
\centering
\includegraphics[width=3in]{figures/Ricker_wavelet.png}
\caption{We use the standard definition of a Ricker (i.e., second derivative of a Gaussian). Image taken from \url{http://subsurfwiki.org}.}
\label{fig:RickerWavelet}
\end{figure}
%%
Originally, if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
Heaviside step function with a very short half duration is defined
for elastic cases to represent the permanent slip on the fault while
in the acoustic case a Gaussian source time function with a similarly
short half duration is defined to physically describe actions within
the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
used, a (pseudo) Dirac delta source time function is defined by default.
Any other source-time function may then be obtained by convolution.
\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
information about the source time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
This feature is only relevant for the solver.
%
\end{description}

>>>>

\vspace{1cm}
\noindent {\bf $\bullet$ Sources setup}
\begin{description}
\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a
(tilted) \texttt{FORCESOLUTION} force point source instead of a \texttt{CMTSOLUTION}
moment-tensor source. When the force source does not fall exactly
at a grid point, the solver interpolates the force between grid points
using Lagrange interpolants. This can be useful e.g. for oil industry
foothills simulations in which the source is a vertical force, normal
force, tilted force, or an impact etc. Note that in the \texttt{FORCESOLUTION}
file, you will need to edit the East, North and vertical components
of an arbitrary (not necessarily unitary, the code will normalize it automatically) direction vector of the force vector;
thus refer to Appendix~\ref{cha:Coordinates} for the orientation
of the reference frame. This vector is made unitary internally in
the solver and thus only its direction matters here; its norm is ignored
and the norm of the force used is the factor force source times the
source-time function.

When using this option, by default the code can locate the force source
anywhere between mesh points in order to honor its exact location; this
is more precise than using the closest GLL mesh point, but it is also a bit slower.
\item [{\texttt{USE\_RICKER\_TIME\_FUNCTION}}] Turn this flag on to use
a Ricker source-time function, i.e., the second derivative of a Gaussian, instead of the source-time functions
set by default to represent a (tilted) \texttt{FORCESOLUTION} force
point source or a \texttt{CMTSOLUTION} moment-tensor source.
Note that we use the standard definition of a Ricker, for a dominant frequency $f_0$:
$\mathrm{Ricker}(t) = (1 - 2 a t^2) e^{-a t^2}$, with $a = \pi^2 f_0^2$,
whose Fourier transform is thus:
$\frac{1}{2} \frac{\sqrt{\pi}\omega^2}{a^{3/2}}e^{-\frac{\omega^2}{4 a}}$
This gives the wavelet of Figure~\ref{fig:RickerWavelet}.
%%
\begin{figure}[htbp]
\centering
\includegraphics[width=3in]{figures/Ricker_wavelet.png}
\caption{We use the standard definition of a Ricker (i.e., second derivative of a Gaussian). Image taken from \url{http://subsurfwiki.org}.}
\label{fig:RickerWavelet}
\end{figure}
%%
Originally, if a \texttt{CMTSOLUTION} moment-tensor source is used, a (pseudo)
Heaviside step function with a very short half duration is defined
for elastic cases to represent the permanent slip on the fault while
in the acoustic case a Gaussian source-time function with a similarly
short half duration is defined to physically describe actions within
the fluid. Otherwise, if a \texttt{FORCESOLUTION} force source is
used, a (pseudo) Dirac delta source-time function is defined by default.
Any other source-time function may then be obtained by convolution.
\item [{\texttt{PRINT\_SOURCE\_TIME\_FUNCTION}}] Turn this flag on to print
information about the source-time function in the file \texttt{OUTPUT\_FILES/plot\_source\_time\_function.txt}.
This feature is only relevant for the solver.
%
\end{description}

### 2022-12-08 doc/USER_MANUAL/05_running_the_solver.tex
updates doc

For any particular earthquake, the \texttt{CMTSOLUTION} file that
represents the point source may be obtained directly from the Harvard
Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.seismology.harvard.edu}.
It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
\begin{figure}[htp]
\begin{centering}
\includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT.jpg}
\par
\end{centering}
\caption{\texttt{CMTSOLUTION} file based on the
format from the Harvard CMT catalog. \textbf{M} is the moment tensor,
$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
magnitude.}
\label{fig:CMTSOLUTION-file}
\end{figure}
%
The \texttt{CMTSOLUTION} file should be edited in the following way:
\begin{itemize}
\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km). Remark: In principle in the international CMTSOLUTION format in geophysics
the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use FORCESOLUTION to describe the source rather than CMTSOLUTION, we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file}, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth). When option USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file} is on, this remark applies to the position of the receivers as well.
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
we recommend setting the source half-duration parameter \texttt{half
duration} equal to zero, which corresponds to simulating a step source-time
function, i.e., a moment-rate function that is a delta function. If
\texttt{half duration} is not set to zero, the code will use a Gaussian
(i.e., a signal with a shape similar to a `smoothed triangle', as
explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
source-time function with half-width \texttt{half duration}. We prefer
to run the solver with \texttt{half duration} set to zero and convolve
the resulting synthetic seismograms in post-processing after the run,
because this way it is easy to use a variety of source-time functions
(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
that the noise generated in the simulation by using a step source
time function may be safely filtered out afterward based upon a convolution
with the desired source time function and/or low-pass filtering. Use
the serial code \texttt{convolve\_source\_timefunction.f90} and the
script \texttt{convolve\_source\_timefunction.csh} for this purpose,
or alternatively use signal-processing software packages such as SAC
\urlwithparentheses{www.llnl.gov/sac}. Type
{\small
\begin{verbatim}
make xconvolve_source_timefunction
\end{verbatim}
}
to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
to the desired half-duration.

\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
or the centroid time of the earthquake. The start time of the simulation
is $t=-2.0*\texttt{half duration}$ (the 2.0 factor is to make sure the moment
rate function is very close to zero when starting the simulation. This avoids spurious high-frequency oscillations.
Also, for acoustic simulations using a Gaussian STF this is set to a factor 3.0, while Ricker source time functions use a factor 1.2).
To convert to absolute time $t_{\mathrm{abs}}$, set
%
\begin{lyxcode}
$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
\end{lyxcode}

@@

If you know the earthquake source in strike/dip/rake format rather
than in \texttt{CMTSOLUTION} format, use the C code \texttt{SPECFEM3D\_GLOBE/utils/strike\_dip\_rake\_to\_CMTSOLUTION.c}
to convert it. The conversion formulas are given for instance in \citet{AkRi80}.
Note that the \citet{AkRi80} convention is slightly different from
the Harvard \texttt{CMTSOLUTION} convention (the sign of some components
is different). The C code outputs both.\newline

Centroid latitude and longitude should be provided in geographical
coordinates. The code converts these coordinates to geocentric coordinates~\citep{DaTr98}.
Of course you may provide your own source representations by designing
your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
file adheres to the Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary
Determination of Earthquakes (PDE) solution performed by the USGS
NEIC, which is used as a seed for the Harvard CMT inversion. The PDE
solution is based upon P waves and often gives the hypocenter of the
earthquake, i.e., the rupture initiation point, whereas the CMT solution
gives the `centroid location', which is the location with dominant
moment release. The PDE solution is not used by our software package
but must be present anyway in the first line of the file.\newline

@@

\noindent The \texttt{FORCESOLUTION} file should be edited in the
following way:
\begin{itemize}
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item Set the \texttt{f0} parameter (the dominant frequency) of the Ricker source time
function (i.e., the second derivative of a Gaussian) when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
in the main parameter file \texttt{DATA/Par\_file}. In case that the
solver uses a (pseudo) Dirac delta source time function to represent
a force point source, a very short duration of five time steps is automatically set by default.
Note that we use the standard definition of a Ricker, for a dominant frequency $f_0$:
$\mathrm{Ricker}(t) = (1 - 2 a t^2) e^{-a t^2}$, with $a = \pi^2 f_0^2$,
whose Fourier transform is thus:
$\frac{1}{2} \frac{\sqrt{\pi}\omega^2}{a^{3/2}}e^{-\frac{\omega^2}{4 a}}$
This gives the wavelet of Figure~\ref{fig:RickerWavelet}.
\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
\item Set the magnitude of the force source.
\item Set the components of a (non-unitary) direction vector for the force
source in the East/North/Vertical basis (see Appendix A for the orientation
of the reference frame).
\end{itemize}
\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
At least one entry (not necessarily the first) must have a zero \texttt{time
shift}, and all the other entries must have non-negative \texttt{time
shift}. Each subevent can have its own latitude, longitude, depth,
\texttt{half duration} and force parameters.

@@

In addition to inbuild source time function, the solver can also use an external source time function defined by the user. This option can be activated by setting \texttt{USE\_EXTERNAL\_SOURCE\_FILE} to true in \texttt{DATA/Par\_file} and by adding the name of the file containing the source time function at the end of \texttt{FORCESOLUTION} or \texttt{CMTSOLUTION} files.
The source time function file must contain the value of the time step on its first line, and then a single column with the amplitude of the source time function for all the time steps. The time step must be exactly the same as that used for the simulation. Note  when the  \texttt{EXTERNAL\_SOURCE\_FILE} is set to false then the line with the external source time function file must not appear in the files  \texttt{FORCESOLUTION} and \texttt{CMTSOLUTION} otherwise the solver will exit with an error. When using an external source file, you can still set up the source location and directivity as in the default case.
In the FORCESOLUTION file: you set "latorUTM", "longorUTM" and "depth" to define the position of your point source. Then if you want to define a directivity, then change the following lines: "component dir vect source E", "component dir vect source N" and "component dir vect source Z\_UP". What you are doing is simply that you define the source position and directivity the same way as in the default case, but in addition you are specifying the path to read in a non-default source time function from an external file.

@@

\item {\it Procedure 2:} useful for other application domains, in which using the absolute $Z$ position of the sources and receivers is more standard than using their depth from the surface
\begin{enumerate}
\item [-] In principle in the international CMTSOLUTION format in geophysics
the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use FORCESOLUTION to describe the source rather than CMTSOLUTION, we provide an option called USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file}, and if so that position is read from CMTSOLUTION in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth). When option USE\_SOURCES\_RECEIVERS\_Z in the \texttt{DATA/Par\_file} is on, this remark applies to the position of the receivers as well.\newline
\item [-] Let's say you want to place two receivers at (x1,y1,z1) and
(x2,y2,z2). Your STATIONS file should then look like:
{\small \begin{verbatim}
BONE  GR  y1  x1  0.00  z1
BONE  GR  y2  x2  0.00  z2
\end{verbatim}}
The option USE\_SOURCES\_RECEIVERS\_Z set to .true. will then discard the elevation and set
burial as the $z$ coordinate.
Third column is Y and Fourth is X due to the latitude/longitude convention.\newline
\end{enumerate}

>>>>

For any particular earthquake, the \texttt{CMTSOLUTION} file that
represents the point source may be obtained directly from the Global
Centroid-Moment Tensor (CMT) web page \urlwithparentheses{www.globalcmt.org}.
It looks like the example shown in Fig. \ref{fig:CMTSOLUTION-file}.
\begin{figure}[htp]
\begin{centering}
\includegraphics[width=0.8\textwidth]{figures/Hollywood_CMT.jpg}
\par
\end{centering}
\caption{\texttt{CMTSOLUTION} file based on the
format from the Global CMT catalog. \textbf{M} is the moment tensor,
$M_{0}${\small {} }is the seismic moment, and $M_{w}$ is the moment
magnitude.}
\label{fig:CMTSOLUTION-file}
\end{figure}
%
The \texttt{CMTSOLUTION} file should be edited in the following way:
\begin{itemize}
\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km). Remark: In principle in the international CMTSOLUTION format in geophysics
the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use \texttt{FORCESOLUTION} to describe the source rather than \texttt{CMTSOLUTION}, we provide an option called \texttt{USE\_SOURCES\_RECEIVERS\_Z} in the \texttt{DATA/Par\_file}, and if so that position is read from \texttt{CMTSOLUTION} in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth). When option \texttt{USE\_SOURCES\_RECEIVERS\_Z} in the \texttt{DATA/Par\_file} is on, this remark applies to the position of the receivers as well.
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item For point-source simulations (see finite sources, page \pageref{To-simulate-a}),
setting the source half-duration parameter \texttt{half duration} equal to zero corresponds to simulating a step source-time
function (Heaviside), i.e., to using a moment-rate function that is a delta function. If
\texttt{half duration} is not set to zero, the code will use a smooth (pseudo) Heaviside source-time function
with a corresponding Gaussian moment-rate function
(i.e., a signal with a shape similar to a `smoothed triangle', as
explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
with half-width \texttt{half duration}.

Often, it is preferable to run the solver with \texttt{half duration} set to zero and convolve
the resulting synthetic seismograms in post-processing after the run,
because this way it is easy to use a variety of source-time functions
(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
that the noise generated in the simulation by using a step source
time function may be safely filtered out afterward based upon a convolution
with the desired source-time function and/or low-pass filtering. Use
the serial code \texttt{convolve\_source\_timefunction.f90} and the
script \texttt{convolve\_source\_timefunction.csh} for this purpose,
or alternatively use signal-processing software packages such as SAC
\urlwithparentheses{www.llnl.gov/sac}. Type
{\small
\begin{verbatim}
make xconvolve_source_timefunction
\end{verbatim}
}
to compile the code and then set the parameter \texttt{hdur} in \texttt{convolve\_source\_timefunction.csh}
to the desired half-duration.

\item The zero time of the simulation corresponds to the center of the triangle/Gaussian,
or the centroid time of the earthquake. The start time of the simulation
is $t=-2.0*\texttt{half duration}$ (the 2.0 factor is to make sure the moment-rate function is very close to zero
when starting the simulation. This avoids spurious high-frequency oscillations.
Also, for acoustic simulations using a Gaussian STF this is set to a factor 3.0, while Ricker source-time functions use a factor 1.2).
To convert to absolute time $t_{\mathrm{abs}}$, set
%
\begin{lyxcode}
$t_{\mathrm{abs}}=t_{\mathrm{pde}}+\texttt{time shift}+t_{\mathrm{synthetic}}$
\end{lyxcode}

@@

If you know the earthquake source in strike/dip/rake format rather
than in \texttt{CMTSOLUTION} format, use the C code \texttt{SPECFEM3D\_GLOBE/utils/strike\_dip\_rake\_to\_CMTSOLUTION.c}
to convert it. The conversion formulas are given for instance in \citet{AkRi80}.
Note that the \citet{AkRi80} convention is slightly different from
the Global/Harvard \texttt{CMTSOLUTION} convention (the sign of some components
is different). The C code outputs both.\newline

Centroid latitude and longitude should be provided in geographical
coordinates. The code converts these coordinates to geocentric coordinates~\citep{DaTr98}.
Of course you may provide your own source representations by designing
your own \texttt{CMTSOLUTION} file. Just make sure that the resulting
file adheres to the Global/Harvard CMT conventions (see Appendix~\ref{cha:Coordinates}).
Note that the first line in the \texttt{CMTSOLUTION} file is the Preliminary
Determination of Earthquakes (PDE) solution performed by the USGS
NEIC, which is used as a seed for the Global/Harvard CMT inversion. The PDE
solution is based upon P waves and often gives the hypocenter of the
earthquake, i.e., the rupture initiation point, whereas the CMT solution
gives the `centroid location', which is the location with dominant
moment release. The PDE solution is not used by our software package
but must be present anyway in the first line of the file.\newline

@@

\noindent The \texttt{FORCESOLUTION} file should be edited in the
following way:
\begin{itemize}
\item The first line is only the header for the force solution, which can
be used as the identifier for the force source.
\item \texttt{time shift:} For a single force source, set
this  parameter equal to $0.0$ (The solver will not run otherwise!); the time
shift parameter would simply apply an overall time shift to the synthetics,
something that can be done in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item \texttt{hdurorf0:} Set the half duration value (s) for step source-time
function, or the dominant frequency (Hz) for Ricker source-time functions.
In case that the source uses a (pseudo) Dirac delta source-time function (i.e., a Gaussian with a zero half duration) to represent
a force point source, a very short half duration of five time steps is automatically set by default.
For a Ricker source-time function, set the dominant frequency value (f0) in Hz.
See the parameter \texttt{source time function:} below for source-time functions.
\item \texttt{latorUTM:} Set the latitude or UTM $x$ coordinate.
\item \texttt{longorUTM:} Set the longitude or UTM $y$ coordinate.
\item \texttt{depth:} Set the depth of the source (in km).
\item \texttt{source time function:} Set the type of source-time function: 0 = Gaussian function, 1 = Ricker function, 2 = Heaviside (step) function, 3 = monochromatic function, 4 = Gaussian function as defined in \citet{Meschede2011}.
When {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on in the main parameter file \texttt{DATA/Par\_file},
it will override this source time function type selection and always use a Ricker wavelet.
Note that we use the standard definition of a Ricker, for a dominant frequency $f_0$:
$\mathrm{Ricker}(t) = (1 - 2 a t^2) e^{-a t^2}$, with $a = \pi^2 f_0^2$,
whose Fourier transform is thus:
$\frac{1}{2} \frac{\sqrt{\pi}\omega^2}{a^{3/2}}e^{-\frac{\omega^2}{4 a}}$
This gives the wavelet of Figure~\ref{fig:RickerWavelet}.
\item \texttt{factor force source:} Set the magnitude of the force source (units in Newton N).
\item \texttt{component dir vect source E:} Set the East component of a direction vector for the force
source. Direction vector is not necessarily a unit vector.
\item \texttt{component dir vect source N:} Set the North component of a direction vector for the force
source.
\item \texttt{component dir vect source Z\_up:} Set the vertical component of a direction vector for the force
source. Sign convention follows the positive upward direction (see Appendix A for the orientation
of the reference frame).
\end{itemize}
\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
At least one entry (not necessarily the first) must have a zero \texttt{time
shift}, and all the other entries must have non-negative \texttt{time shift}.
Each subevent can have its own set of parameters \texttt{latitude}, \texttt{longitude}, \texttt{depth},
\texttt{half duration} etc.\newline

@@

In addition to inbuild source-time functions, the solver can also use an external source-time function defined by the user. This option can be activated by setting \texttt{USE\_EXTERNAL\_SOURCE\_FILE} to \texttt{.true.} in \texttt{DATA/Par\_file} and by adding the name of the file containing the source-time function at the end of \texttt{FORCESOLUTION} or \texttt{CMTSOLUTION} files.
The source-time function file must contain a single column with the amplitudes of the source-time function for all the time steps. The time step must be exactly the same as that used for the simulation. When the flag is set to \texttt{.false.}, then the line with the external source-time function file must not appear in the files \texttt{FORCESOLUTION} and \texttt{CMTSOLUTION}, otherwise the solver will exit with an error. When using an external source file, you can still set up the source location and directivity as in the default case.
In the \texttt{FORCESOLUTION} file: you set "latorUTM", "longorUTM" and "depth" to define the position of your point source. Then if you want to define a directivity, change the following lines: "component dir vect source E", "component dir vect source N" and "component dir vect source Z\_UP". What you are doing is simply that you define the source position and directivity the same way as in the default case, but in addition you are specifying the path to read in a non-default source-time function from an external file.

@@

\item {\it Procedure 2:} useful for other application domains, in which using the absolute $Z$ position of the sources and receivers is more standard than using their depth from the surface
\begin{enumerate}
\item [-] In principle in the international CMTSOLUTION format in geophysics
the depth is given in kilometers; however for users in other fields (non-destructive testing, medical imaging, near-surface studies...) who may prefer to give the position of the source (rather than its depth from the surface), or for people who use \texttt{FORCESOLUTION} to describe the source rather than \texttt{CMTSOLUTION}, we provide an option called \texttt{USE\_SOURCES\_RECEIVERS\_Z} in the \texttt{DATA/Par\_file}, and if so that position is read from \texttt{CMTSOLUTION} in meters rather than kilometers (and again, it is then the true position in the mesh, not the depth). When option \texttt{USE\_SOURCES\_RECEIVERS\_Z} in the \texttt{DATA/Par\_file} is on, this remark applies to the position of the receivers as well.\newline
\item [-] Let's say you want to place two receivers at (x1,y1,z1) and
(x2,y2,z2). Your STATIONS file should then look like:
{\small \begin{verbatim}
BONE  GR  y1  x1  0.00  z1
BONE  GR  y2  x2  0.00  z2
\end{verbatim}}
The option USE\_SOURCES\_RECEIVERS\_Z set to .true. will then discard the elevation and set
burial as the $z$ coordinate.
Third column is Y and Fourth is X due to the latitude/longitude convention.\newline
\end{enumerate}

### 2022-12-08 doc/USER_MANUAL/14_post_processing.tex
updates doc

For such comparisons, the following steps are recommended:
\begin{enumerate}
\item Make sure that both synthetic and observed seismograms have the correct
station/event and timing information.
\item Convolve synthetic seismograms with a source time function with the
half duration specified in the \texttt{CMTSOLUTION} file, provided,
as recommended, you used a zero half duration in the SEM simulations.
\item Resample both observed and synthetic seismograms to a common sampling
rate.
\item Cut the records using the same window.
\item Remove the trend and mean from the records and taper them.
\item Remove the instrument response from the observed seismograms (recommended)
or convolve the synthetic seismograms with the instrument response.
\item Make sure that you apply the same filters to both observed and synthetic
seismograms. Preferably, avoid filtering your records more than once.
\item Now, you are ready to compare your synthetic and observed seismograms.
\end{enumerate}
\noindent We generally use the following scripts provided in the \texttt{utils/seis\_process/}
directory:

@@

More options are available for this script, such as adding time shift
to the origin time of the synthetics, convolving the synthetics with
a triangular source time function with a given half duration, etc.
Type \texttt{process\_syn.pl} without any argument for a detailed
usage.\newline

>>>>

For such comparisons, the following steps are recommended:
\begin{enumerate}
\item Make sure that both synthetic and observed seismograms have the correct
station/event and timing information.
\item Convolve synthetic seismograms with a source-time function with the
half duration specified in the \texttt{CMTSOLUTION} file, provided,
as recommended, you used a zero half duration in the SEM simulations.
\item Resample both observed and synthetic seismograms to a common sampling
rate.
\item Cut the records using the same window.
\item Remove the trend and mean from the records and taper them.
\item Remove the instrument response from the observed seismograms (recommended)
or convolve the synthetic seismograms with the instrument response.
\item Make sure that you apply the same filters to both observed and synthetic
seismograms. Preferably, avoid filtering your records more than once.
\item Now, you are ready to compare your synthetic and observed seismograms.
\end{enumerate}
\noindent We generally use the following scripts provided in the \texttt{utils/seis\_process/}
directory:

@@

More options are available for this script, such as adding time shift
to the origin time of the synthetics, convolving the synthetics with
a triangular source-time function with a given half duration, etc.
Type \texttt{process\_syn.pl} without any argument for a detailed
usage.\newline
